---
title: On Average, are Diamonds Certified by HRD Priced Higher than those Certified
  by Other Organizations?
author: "Prasanna K R"
date: "August 31, 2015"
output: 
  pdf_document: 
    keep_tex: yes
    latex_engine: lualatex
    number_sections: yes
  html_document: 
    keep_md: yes
    number_sections: yes
---


```{r setup, echo=FALSE , include=FALSE}
knitr::opts_chunk$set(cache=TRUE, out.width='450px' , dpi=200 )  
options(width=80 , scipen=10000)
```

```{r libraries, include=FALSE}
library(ggplot2)
```

# The Hypotheses
Let us state our Hypotheses:

* <B>Null Hypothesis (H~0~)</B> : On average, there is no difference in diamond's pricing, between HRD or any other organization
* <B>Alternate Hypothesis (H~A~)</B>: On average, HRD certified diamonds are priced higher (one-sided test)

Mathematically, the hypothese are expressed below:

* H~0~: μ~diff~ = 0
* H~A~: μ~diff~ > 0

# The Data
```{r hypDataSubset, echo=FALSE}
diamonds <- read.csv( "Data.csv")
HRD_Diamonds <- subset( diamonds , diamonds$Certification.Body == "HRD")
Non_HRD_Diamonds <- subset( diamonds , diamonds$Certification.Body != "HRD" )
```

We separate the data sets - diamonds certified by HRD and other diamonds. There are ```r nrow(HRD_Diamonds)``` observations of HRD Diamonds and ```r nrow( Non_HRD_Diamonds )``` non-HRD Certified Diamonds.

# Central Limit Theorm: Checking the Conditions for Hypothesis Testing for Paired Data
The conditions for hypothesis testing:

* <B>Independence:</B> Sampled observations must be independent. Random sample must be collected with replacement because the size of HRD Certified diamond is small
* <B>Sample Size / Skew: </B> The no of elements must be more than 30.

We select a size of <code>40</code> with replacement.

```{r samplingHyp, echo=FALSE}
ndiff <- 40
set.seed(425)
HRD_Sample <- sample( x = HRD_Diamonds$Price.in.Singapore.. , size = ndiff , replace = TRUE )
Non_HRD_Sample <- sample( x = Non_HRD_Diamonds$Price.in.Singapore.. , size = ndiff , replace = TRUE )
```

```{r vizualize, echo=FALSE}
xVar <- c(HRD_Sample , Non_HRD_Sample )
colorVar <- as.factor(c(rep(1, ndiff), rep(2, ndiff ) ) )
sampleDensityDf <- data.frame( xVar ,  colorVar )

ggplot( sampleDensityDf , aes(x = xVar, fill = colorVar) ) + 
  geom_density( alpha = .2 ) +
  scale_x_continuous( "Price of Diamond in SGD" ) +
  scale_fill_discrete( 
    name = "Sample" , labels=c( "HRD Certified Sample", "Non HRD Certified Sample" ) ) +
  scale_y_continuous( "Density" )

ggplot( sampleDensityDf , aes( colorVar , xVar ) ) + 
  geom_boxplot( aes( fill = colorVar ) ) + 
  scale_y_continuous( "Price of Diamond in SGD" ) +
  scale_fill_discrete( 
    name = "Sample" , labels=c( "HRD Certified Sample", "Non HRD Certified Sample" ) ) +
  scale_x_discrete( "Sample" , labels=c( "HRD Certified Sample", "Non HRD Certified Sample" )  )

```

# The Test Statistic
```{r calculTestStatistic, echo=FALSE}
standardError <- function( x ) {
  sd( x )/sqrt( length( x ) )
}

Diff = HRD_Sample - Non_HRD_Sample
xBar <- mean(Diff)
```
```{r calculTestStatistic1}
zScore <- xBar / standardError(Diff)
zScore
# Calculating p-value
# 1-pnorm() because we are doing a one-sided test - greater than
pValue <- 1-pnorm( zScore ) 
pValue

```

# Null Hypothesis (H~0~) is Rejected
The **Null Hypothesis (H~0~)** is rejected because the <code>pValue</code> is lesser than the significance value of <code>0.05</code>.

